---
title: "Portfolio 7"
author: "Conor Lacey"
date: "2023-04-13"
output: github_document
---

```{r libraries}
suppressWarnings(library(tidyverse))
```

### Introduction

In this portfolio I am going to be trying out gganimate! In presenting for SAM I felt it would be helpful to animate some Bayesian concepts. More specifically I hope to animate how prior information updates your posterior distribution so as to get closer to the truth. 

### gganimate Libraries

First let's load the appropriate libraries

```{r gganimate libraries}
library(gganimate)
library(transformr)
```

gganimate is the primary package that will make it user friendly to code a moving graph, however I also need to the transformr package because this allows for interpolation between data points (i.e., this will make the graph movement smooth)

### Normal Distribution

Now let's make normal distribution centered at 0 and with an SD of 1. 

```{r normal distribution}
x <- seq(-4,4, length.out = 1e4)
y <- dnorm(x, mean = 0, sd = 1)
dat <- tibble(x = x, y = y)

dat %>% ggplot(aes(x = x, y = y)) +
  geom_line(alpha=0.5, linewidth = 2, color = "#4169E1") +
  labs(title = "Normal Distribution",
       y = "Density")
```

We should also make the normal distribution that we would like the previous above distribution to shift to. 

```{r new normal}
x <- seq(-4,4, length.out = 1e4)
y <- dnorm(x, mean = 2, sd = .5)
dat <- tibble(x = x, y = y)

dat %>% ggplot(aes(x = x, y = y)) +
  geom_line(alpha=0.5, linewidth = 2, color = "#4169E1") +
  labs(title = "Normal Distribution",
       y = "Density")
```

### gganimate (A rookie mistake)

Alright well now this should be simple right? Let's just code these two graphs so that there is a transition between the two using gganimate.

```{r gganimate rookie}
prior <- tibble(x = seq(-4,4, length.out = 1e4),
                y = dnorm(x, mean = 0, sd = 1),
                group =1)

post <-  tibble(x = seq(-4,4, length.out = 1e4),
                y = dnorm(x, mean = 2, sd = .5),
                group =2)

prior_post <- bind_rows(prior,post)

prior_post %>% ggplot(aes(x=x, y=y)) + 
  geom_line(alpha=0.5, linewidth = 2, color = "#4169E1") +
  #Animation
  labs (x = "dMACS",
        y = "Density") + 
  transition_time(group) + 
  ease_aes('linear')
```

Well this isn't quite what we are looking for... 
I mean yes, it's changing from the prior to the posterior distribution, but it's not shifting. 

The root of the problem comes from the group variable. Yes, these distributions are different types and hence should be put in different groups. However, the way gganimate works is that it is using this grouping variable as its indicator for what frames to shift between. If we only have two frames, then it's going to shift between these two groups directly, which won't look very smooth. Therefore, to make this smoother we need more frames or groups of normal distributions that are incrementally increasing to a mean of 2 and decreasing to a sd of 0.5. To simplify the coding of this further I will now call the group variable frames. 

### gganimate (The eliquent way)

```{r gganimte eliquent}
a <- seq(0, 2, length = 100)
b <- seq(1, .5, length = 100)
```

The a and b vectors will specify how we want the distribution to incrementally change.

Now let's create a for loop to create these distributions frame by frame. We will make it so that there are 100 frames of normal distributions that are increasing from a mean of 0 to a mean 2 and decreasing from a sd of 1 to an sd of 0.5. 

```{r for loop}
prior_post<- tibble(x = 0, y = 0, frame = 0)
for (i in 1:100){
x<-seq(-4,4, length.out = 1e4)
y<-dnorm(x, a[i], b[i])
y<-y/max(y)*(.5)
dat <- tibble(x = x, y = y, frame = i)
prior_post <- bind_rows(prior_post, dat)
}
prior_post <- prior_post[-1,]
```

Lastly, let's animate it!

```{r animate!}
prior_post %>% ggplot(aes(x=x, y=y)) + 
  geom_line(alpha=0.5, linewidth = 2, color = "#4169E1") +
  #Animation
  labs (x = "dMACS",
        y = "Density") + 
  transition_time(frame) + 
  ease_aes('linear')
```

Nice! This is very smooth. Now how about we add some more information on this graph. Specifically let's add a marker for where the true value is in the population as well as the percent of prior knowledge to represent how increasing your prior knowledge gets to you closer to the truth. 

```{r more info}
prior_post %>% ggplot(aes(x=x, y=y)) + 
  geom_line(alpha=0.5, linewidth = 2, color = "#4169E1") +
  geom_vline(xintercept = 2) +
  #Animation
  labs (title = "Amount of Prior Knowledge = {frame_time}%",
        subtitle ="                                                                                                True Value",
        x = "dMACS",
        y = "Density") + 
  transition_time(frame) + 
  ease_aes('linear')
```

And there we have it! 
